---
title: "Predict422-CharityProject Part 1"
author: "Artur Mrozowski"
date: "April 19, 2017"
output: pdf_document
---

Exercises 1. Read Data from CSV File 
(a) Read data from csv file

```{r}
charityData = read.csv(file.choose(),na.strings=c("NA"," "))
```
(c)
```{r}
charityData$DONR = as.factor(charityData$DONR)
charityData$HOME = as.factor(charityData$HOME)
charityData$HINC = as.factor(charityData$HINC)
```
Following variable converted to factors. GENDER and RFA_96 converted due to default setting stringsAsFactors = TRUE 
```{r}
str(charityData)
```
2.Data Quality Check 
```{r}
dim(charityData)      # dimensions of data
names(charityData)    # variable names
#str(charityData)      # one form of summary of data
  # another form of summary
```
(b) We can see that HOME, HINC and GENDER columns contain some missing values.
```{r}
## Check for Missing Values
which(sapply(charityData,anyNA))

# Missing values identified in HINC, GENDER, and RFA_96
# Get counts of missing values for each variable
table(charityData$HOME,useNA="ifany")
table(charityData$HINC,useNA="ifany")
table(charityData$GENDER,useNA="ifany")
```
(c) Missing values could be a problem. Replacing the missing values with mean could be one solution.



3(a)
Histogram of the response variable DAMT
(first with 0s included, second with 0s dropped)
```{r}
hist(charityData$DAMT,col="blue",breaks=20,xlab="DAMT",main="")
hist(charityData$DAMT[charityData$DAMT > 0],col="blue",breaks=20,xlab="DAMT > 0",main="")
```

```{r}
hist(charityData$AGE)
hist(charityData$MEDAGE)
hist(charityData$MEDPPH)
hist(charityData$MEDHVAL)
hist(log(charityData$MEDHVAL))
hist(charityData$MEDINC)
hist(charityData$MEDEDUC)
hist(charityData$NUMPROM)
hist(charityData$NUMPRM12)
hist(charityData$RAMNTALL)
hist(charityData$NGIFTALL)
hist(charityData$MAXRAMNT)
hist(charityData$LASTGIFT)
hist(charityData$TDON)
```

```{r}
table(charityData$DONR)
barplot(table(charityData$DONR),xlab="DONR")

```
(b)
Boxplot of DAMT amount by categories for GENDER
```{r}
plot(charityData$GENDER,charityData$DAMT,xlab="Gender",ylab="Donation ($)")

```
Plot DAMT against a quantitative predictor variable
```{r}


plot(charityData$AGE,charityData$DAMT,xlab="Age",ylab="Donation ($)")
lm_age = lm(DAMT ~ AGE, data=charityData)
abline(lm_age,col="red")

```

Let's see which variable have the greates potential for the regression. 
```{r}
charity_damt<-subset(charityData,DAMT > 0,select=c(DAMT,AGE,MEDAGE,MEDPPH,MEDHVAL,MEDINC,MEDEDUC,NUMPROM
                                         ,NUMPRM12,RAMNTALL,NGIFTALL,MAXRAMNT,LASTGIFT,TDON))
cor(charity_damt)

```
LASTGIFT and MAXRAMNT have the greatest potential for the regression problem.

(c)
Boxplot of AGE by DONR status

```{r}
plot(charityData$DONR,charityData$AGE,xlab="DONR",ylab="AGE")
```
Wrong" Way. There are more females than males in the dataset as a whole.
```{r}
barplot(table(charityData$GENDER[charityData$DONR == 1]),xlab="GENDER",main="Barplot of GENDER for DONR = 1")
```
Right Way
# A mosaic plot is obtained when we plot one factor variable against another. The
# mosaic plot represents the counts as proportions to the whole. A deviation in
# overall proportion of females donating compared to males donating is meaningful
# whereas the absolute count of females donating compared to males donating was not.

```{r}
plot(charityData$DONR,charityData$GENDER,xlab="DONR",ylab="GENDER",main="Mosaic Plot")
# Or
plot(charityData$GENDER,charityData$DONR,xlab="GENDER",ylab="DONR",main="Mosaic Plot")
```
# These graphs show that M/F doesn't show any difference in DONR status.
```{r echo=FALSE}
separateRFA = function(xData,varName)
{
  bytes = c("R","F","A")
  newVarNames = paste(varName,bytes, sep="_")
  
  for (ii in 1:length(bytes)) # Loop over 1 to 3 (corresponding to R, F, and A)
  {
    # Find the unique values for current byte
    byteVals = unique(substr(levels(xData[,varName]),ii,ii))
    
    for (jj in 1:length(byteVals)) # Loop over unique byte values
    {
      rowIdx = substr(xData[,varName],ii,ii) == byteVals[jj]
      xData[rowIdx,newVarNames[ii]] = byteVals[jj]
    }
    
    xData[,newVarNames[ii]] = factor(xData[,newVarNames[ii]])
  }
  
  return(xData)
}

```

```{r}
charityData = separateRFA(charityData,"RFA_96")
# Check the results
table(charityData$RFA_96,charityData$RFA_96_R)
table(charityData$RFA_96,charityData$RFA_96_F)
table(charityData$RFA_96,charityData$RFA_96_A)

```

Excercise 4
First we will subset the charity data set and exclude all the categorical variables.Not sure how to handle mix of continous and categorical variables. We only include DAMT values greater than 0.

```{r}
charity_sub<-subset(charityData,DAMT > 0,select=c(AGE,MEDAGE,MEDPPH,MEDHVAL,MEDINC,MEDEDUC,NUMPROM
                                         ,NUMPRM12,RAMNTALL,NGIFTALL,MAXRAMNT,LASTGIFT,TDON))
```
Let's look at mean an variance of the variables. It 
```{r}
sapply(charity_sub,mean)
sapply(charity_sub, var)
```
We now perform principal component analysis using prcomp function.
```{r}
pr.out<-prcomp(charity_sub,scale=TRUE)
pr.out$sdev
```
The center and scale varialbes correspond to meand and standard deviations of variable prior to PCA.
```{r}
pr.out$center
pr.out$scale

```

```{r}
pr.out$rotation
```

```{r}
biplot(pr.out,scale=0)


```


```{r}
pr.out$sdev

pr.var<-pr.out$sdev^2
pve=pr.var/sum(pr.var)
pve
```

```{r}
plot(pve , xlab=" Principal Component ", ylab=" Proportion ofVariance Explained ", ylim=c(0,1),type="b" )
plot(cumsum (pve ), xlab=" Principal Component ", ylab ="Cumulative Proportion of Variance Explained ", ylim=c(0,1),type="b" )
```
Looks like first 6 principal components explain 80% of the variance in 
```{r}

```

```{r}

```


```{r}

```

```{r}

```

```{r}

```

```{r}

```



