---
title: "Predict422-CharityProject Part4"
author: "Artur Mrozowski"
date: "May 31, 2017"
output: pdf_document
---



```{r}
# Load packages required for this code.
library(glmnet)

```

1. Import Data
a. Read the data into R from the CSV file valSample.csv the same as you did in Part 1 Exercise 1.
b. Subset the data to only those observations where DONR = 1. This set of observations will be the data used
for this assignment (Part 2, The Regression Problem). Name this dataset valData.
```{r}
inPath = file.path("C:\\playground\\Predict422\\Project\\Part4\\")

valDataIn = read.csv(file.path(inPath,"valSample.csv"),na.strings=c("NA"," "))
#valData = valDataIn[valDataIn$DONR == "1",]
valData=valDataIn
rm(valDataIn)

```


```{r}
# Convert categorical variables to factors
# This is highly recommended so that R treats the variables appropriately.
# The lm() method in R can handle a factor variable without us needing to convert 
# the factor to binary dummy variable(s).
valData$DONR = as.factor(valData$DONR)
valData$HOME = as.factor(valData$HOME)
valData$HINC = as.factor(valData$HINC)
```
2. Predictions on Validation Set
a. Review the data preparation steps you took in Part 2 of the project. Apply those same data preparation steps
to valData.
```{r}

codePath = file.path("C:\\playground\\Predict422\\Project\\Part4\\")
source(file.path(codePath,"DataPreparation.R"))

valDataPart2 = processPart2(valData)
```

b. Using the model you chose from Part 2 (as trained on the Regression Training Set from Part 2), predict
DAMT on the data coming from Step 2a.
```{r}
# Note that RFA_96_A for valData does not include the level "B". I had to do some 
# investigating to track down an error that originated from this fact. Therefore, we
# will add the level so that we don't have problems with making predictions.
levels(valDataPart2$RFA_96_A) = c(levels(valDataPart2$RFA_96_A), "B")


```


```{r}
modelPath = file.path("C:\\playground\\Predict422\\Project\\Part4\\")
load(file.path(modelPath,"modelPart2.RData"))
```


```{r}
valData$DAMT.Pred = as.numeric(predict(modelPart2,newdata=valDataPart2))

# Check the predictions as a sanity check
hist(valData$DAMT.Pred,xlab="DAMT",main="Validation Set",col="gray",breaks=50)
par(pty="s")
plot(valData$DAMT,valData$DAMT.Pred,xlab="Target DAMT",ylab="Predicted DAMT",
     main="Validation Set")
abline(0,1,col="red")
par(pty="m")
```

c. Review the data preparation steps you took in Part 3 of the project. Apply those same data preparation steps
to valData.
```{r}
valDataPart3 = processPart3(valData)
levels(valDataPart3$RFA_96_A) = c(levels(valDataPart3$RFA_96_A), "B")
```

d. Using the model you chose from Part 3 (as trained on the Classification Training Set from Part 3), predict
DONR and PDONR on the data coming from Step 2c.
```{r}
load(file.path(modelPath,"modelPart3.RData"))
```


```{r}
assignClass = function(probVals,threshVal)
{
  predVals = rep(0,length(probVals))
  predVals[probVals > threshVal] = 1
  predVals = factor(predVals)
  
  return(predVals)
}
```


```{r}
# Further note that for a logistic regression model, the probabilities (PDONR) come 
# from predict.glm and the classifications (DONR) come from applying the optimal threshold.
# Each predict method should have some means of obtaining the probabilities. You
# will have to check the documentation for the type of model you are using to 
# determine the appropriate syntax.
valData$PDONR.Pred = predict(modelPart3,newdata=valDataPart3,type="response")
valData$DONR.Pred = assignClass(valData$PDONR.Pred,optThreshPart3)
```


```{r}
# Check the predictions as a sanity check
table(valData$DONR,valData$DONR.Pred,dnn=c("Target","Predicted"))
hist(valData$PDONR.Pred,xlab="P(DONR=1)",main="Validation Set",col="gray",breaks=50,
     xlim=c(0,1))
plot(valData$DONR,valData$PDONR.Pred,xlab="Target DONR Value",
     ylab="Predicted P(DONR=1)",main="Validation Set")
```

3. Mailing List Selection
```{r}
source(file.path(codePath,"RankedDonorOutput.R"))
```

a. The mailing list selection strategy illustrated in the sample code requires you to choose a score to rank and
select a cutoff to use on that score. Evaluate this strategy by ranking various scores and calculating the
profit obtained on the validation dataset. Scores that you might consider using include the predicted values
of DONR, PDONR, and EXAMT. Summarize your findings with tables and figures as appropriate.
```{r}
# Rank donors by PDONR.Pred
numBins = 10
out1 = outputForRankedDonors(numBins,rankVar="PDONR.Pred",dataToRank=valData)
print(out1$Donor.Table)
print(out1$Mailing.Table)
```


```{r}
# Rank donors by EXAMT.Pred (expected donation amount)
# EXAMT.Pred = PDONR.Pred * DAMT.Pred 
# (likelihood of donation * predicted donation amount)
valData$EXAMT.Pred = valData$PDONR.Pred * valData$DAMT.Pred
out2 = outputForRankedDonors(numBins,rankVar="EXAMT.Pred",dataToRank=valData)
print(out2$Donor.Table)
print(out2$Mailing.Table)
```


```{r}
# Rank donors by DAMT.Pred (predicted donation amount)
out3 = outputForRankedDonors(numBins,rankVar="DAMT.Pred",dataToRank=valData)
print(out3$Donor.Table)
print(out3$Mailing.Table)
```


```{r}
# Calculate percentiles of breakVals for each profile using the empircal CDF function.
fn1 = ecdf(out1$breakVals)
fn2 = ecdf(out2$breakVals)
fn3 = ecdf(out3$breakVals)
yLimits = c(-500+1000*floor(min(c(
              out1$Mailing.Table$Total.Profit,
              out2$Mailing.Table$Total.Profit,
              out3$Mailing.Table$Total.Profit
            ))/1000),
            500+1000*ceiling(max(c(
              out1$Mailing.Table$Total.Profit,
              out2$Mailing.Table$Total.Profit,
              out3$Mailing.Table$Total.Profit
            ))/1000))
plot(fn1(out1$breakVals)[-1],out1$Mailing.Table$Total.Profit,type='b',col="blue",
     xlab="% Mailed",ylab="Profit ($)",main="Profit Profiles",xlim=c(0,1),ylim=yLimits)
lines(fn2(out2$breakVals)[-1],out2$Mailing.Table$Total.Profit,col="red")
points(fn2(out2$breakVals)[-1],out2$Mailing.Table$Total.Profit,col="red",pch=16)
lines(fn3(out3$breakVals)[-1],out3$Mailing.Table$Total.Profit,col="green")
points(fn3(out3$breakVals)[-1],out3$Mailing.Table$Total.Profit,col="green",pch=16)
abline(h=0,lty=2)
legend(x="bottomleft",legend=c("PDONR","EXAMT","DAMT","Break Even"),
       col=c("blue","red","green","black"),
       lty=c(1,1,1,2),pch=c(1,16,16,NA))
```


```{r}
cutOff = out3$breakVals[numBins+1-2]
valMailList = data.frame(ID=valData$ID[valData$DAMT.Pred >= cutOff])
length(valMailList$ID)
```
4. Predictions on Test Set
In this exercise, you will make predictions on the Test Set data provided in testSample.csv. You will then
select individuals from the Test Set to be mailed in the upcoming charity mailing campaign.


a. Repeat Exercise 1 of this assignment applied to the data in testSample.csv.
```{r}
testData = read.csv(file.path(inPath,"testSample.csv"),na.strings=c("NA"," "))

testData$HOME = as.factor(testData$HOME)
testData$HINC = as.factor(testData$HINC)
```

b. Repeat Exercise 2 of this assignment applied to the data in testSample.csv.
```{r}
# Note: The model.matrix method will not allow us to use a dataframe with "missing" 
# columns. Therefore, we add dummy DAMT and DONR columns to testData.
testData$DAMT = -1
testData$DONR = -1
```


```{r}
## Apply the Part 2 data processing steps to testData
testDataPart2 = processPart2(testData)
levels(testDataPart2$RFA_96_A) = c(levels(testDataPart2$RFA_96_A), "B")
```


```{r}
## Predict DAMT for testData using your chosen model from Part 2
# Note that the model I am using is a glmnet model.
x = model.matrix(DAMT ~ .-ID,data=testDataPart2)[,-1]
testData$DAMT.Pred = as.numeric(predict(modelPart2,newdata = testDataPart2))
```


```{r}
# Check the predictions as a sanity check
summary(testData$DAMT.Pred)
```


```{r}
## Apply the Part 3 data processing steps to valData
testDataPart3 = processPart3(testData)
levels(testDataPart3$RFA_96_A) = c(levels(testDataPart3$RFA_96_A), "B")

## Predict DONR and PDONR for valData using your chosen model from Part 3
# Note that the model I am using is a glm model.
testData$PDONR.Pred = predict(modelPart3,newdata=testDataPart3,type="response")
testData$DONR.Pred = assignClass(testData$PDONR.Pred,optThreshPart3)

# Check the predictions as a sanity check
table(testData$DONR.Pred)
summary(testData$PDONR.Pred)
```

c. Write your predictions out to a CSV file called projectPredictionsTEST.csv. This CSV file
should contain the following columns: ID, DONR, PDONR, and DAMT.
```{r}
## Part C - Write Test Set Predictions to CSV File
# Name the columns in the CSV file ID, DONR, PDONR, DAMT
testPredOut = data.frame(ID = testData$ID,
                         DONR = testData$DONR.Pred,
                         PDONR = testData$PDONR.Pred,
                         DAMT = testData$DAMT.Pred)

outPath = file.path("C:\\playground\\Predict422\\Project\\Part4\\")

write.csv(testPredOut,file=file.path(outPath,"projectPredictionsTEST.csv"),row.names=FALSE)
```

d. Apply the mailing list selection strategy that you chose in Exercise 3b to the Test Set.
```{r}
# Use cutoff selected above.
testMailList = data.frame(ID=testData$ID[testData$DAMT.Pred >= cutOff])
length(testMailList$ID)
```
e. Write the ID numbers of individuals selected for the mailing list to a CSV file called
projectListTEST.csv. This CSV file needs only a single column: ID.

```{r}
write.csv(testMailList,file=file.path(outPath,"projectListTEST.csv"),row.names=FALSE)
```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```


```{r}

```

