---
title: "Predic422-CharityProject Part 3"
author: "Artur Mrozowski"
date: "May 28, 2017"
output: pdf_document
---

Load packages required for this code.

```{r}
# Load packages required for this code.
library(pROC)
library(lift)
library(MASS)
library(rpart)
library(caret)
```

 Exercise 1
Read Data from CSV File
```{r}
charityData = read.csv(file.choose(),na.strings=c("NA"," "))
```
Convert categorical variables to factors

```{r}
charityData$DONR = as.factor(charityData$DONR)
charityData$HOME = as.factor(charityData$HOME)
charityData$HINC = as.factor(charityData$HINC)

```
Rename the dataset to classData for clarity.
Remove charityData from R session environment
```{r}
classData = charityData
rm(charityData)
```


```{r}
## Check for Missing Values
which(sapply(classData,anyNA))
```


```{r}
# HOME - Make a level 0 and code missing values as 0
levels(classData$HOME) = c(levels(classData$HOME),"0")
classData$HOME[is.na(classData$HOME)] = "0"
table(classData$HOME,useNA="ifany")
```


```{r}
# HINC - Make a level 0 and code missing values as 0
levels(classData$HINC) = c(levels(classData$HINC),"0")
classData$HINC[is.na(classData$HINC)] = "0"
table(classData$HINC,useNA="ifany")
```


```{r}
# GENDER - Assign A, J, and NA to category U
idxMF = classData$GENDER %in% c("M","F")
classData$GENDER[!idxMF] = "U"
classData$GENDER = factor(classData$GENDER)
table(classData$GENDER)
```
Part B - Derived or Transformed Variables(Optional)


Part C - Re-categorize Variables
```{r}
# Separate RFA Values (R = recency, F = frequency, A = amount)
separateRFA = function(xData,varName)
{
  bytes = c("R","F","A")
  newVarNames = paste(varName,bytes, sep="_")
  
  for (ii in 1:length(bytes)) # Loop over 1 to 3 (corresponding to R, F, and A)
  {
    # Find the unique values for current byte
    byteVals = unique(substr(levels(xData[,varName]),ii,ii))
    
    for (jj in 1:length(byteVals)) # Loop over unique byte values
    {
      rowIdx = substr(xData[,varName],ii,ii) == byteVals[jj]
      xData[rowIdx,newVarNames[ii]] = byteVals[jj]
    }
    
    xData[,newVarNames[ii]] = factor(xData[,newVarNames[ii]])
  }
  
  return(xData)
}
```


```{r}
# Apply separateRFA to the variables RFA_96 and check results.

classData = separateRFA(classData,"RFA_96")
#table(classData$RFA_96,classData$RFA_96_R)
#table(classData$RFA_96,classData$RFA_96_F)
#table(classData$RFA_96,classData$RFA_96_A)
```

Part D - Drop Variables
```{r}
dropIdx = which(names(classData) %in% c("DAMT","RFA_96"))

# Drop the variables indicated by dropIdx.
classData2 = classData[,-dropIdx]
names(classData2)   # check that the result is as expected
```
Exercise 3
Dataset Partitioning

```{r}
# Specify the fraction of data to use in the hold-out test.
testFraction = 0.25   
set.seed(123)
```


```{r}
# Sample training subset indices.

trainIdx = sample(nrow(classData2),size=(1-testFraction)*nrow(classData2),
                  replace=FALSE)
```
Exercise 4
Model Fitting
```{r}
glm.fit=glm(DONR~ AGE+MEDAGE+MEDHVAL+MEDINC+MEDEDUC+NUMPROM+MAXRAMNT +MEDINC+MEDEDUC+ NUMPROM+NUMPRM12+RAMNTALL+NGIFTALL+MAXRAMNT+TDON+LASTGIFT+RFA_96_R+RFA_96_F+RFA_96_A, data=classData2,subset=trainIdx,family=binomial )
backwards=step(glm.fit,trace=0)
formula(backwards)
summary(backwards)
```

Part A - Simple Logistic Regression

One of the variables with considerable significance is RFA_96_R. I will now fit logistic  regression using that variable. 
```{r}

#modelA1 = glm(DONR ~ MAXRAMNT,data=classData2,subset=trainIdx,family=binomial)
modelA1 = glm(DONR ~ RFA_96_F,data=classData2,subset=trainIdx,family=binomial)
summary(modelA1)
par(mfrow=c(2,2))
plot(modelA1)
par(mfrow=c(1,1))
```


```{r}
trnProbsA1 = predict(modelA1,type="response")
hist(trnProbsA1,col="gray")   # Note that scores are distributed around 0.05.
hist(trnProbsA1,col="gray",xlim=c(0,1))   # Rescale to make obvious.

```


```{r}
# Classification: ROC Curve for Model A1 - Use methods from pROC package.
rocA1 = roc(response=classData2$DONR[trainIdx],predictor=trnProbsA1)
par(pty="s")  # sets plotting region to a square, useful for ROC curves
# Use par(pty="m") to return to default of rectangular plotting region.
plot(rocA1,col="blue",
     main=paste("ROC curve for Model A1\nAUC = ",round(rocA1$auc,digits=3),sep=""))
par(pty="m")
# Classification: Determine "optimal" threshold.
dist01 = sqrt((rocA1$specificities-1)^2 + (rocA1$sensitivities-1)^2)
optIdxA1 = which.min(dist01)  # index corresponding to minimum distance
threshA1 = rocA1$thresholds[optIdxA1]  # threshold corresponding to min. distance
points(rocA1$specificities[optIdxA1],rocA1$sensitivities[optIdxA1],col="red",pch=7)
```


```{r}

```

Note that the `echo = FALSE` parameter was added to the code chunk to prevent printing of the R code that generated the plot.
